QPLIB: A Library of Quadratic Programming Instances

### 2017-01-27 format.tex
Add files via upload

Following the notation of \S~\ref{sec:QPbasic}, a generic Quadratic Program (QP) can be written as follows:
%
\begin{align*}
 \min \;\;
 & x^\top Q^0 x + b^0 x + f \\
 %
 & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn  \\
 %
 & x_j \in \mathbb{Z} & j \in \Mcz
\end{align*}

where:

\begin{itemize}
 \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
 %
 \item $x = [x_j]_{j = 1}^n \in \RR^n$ is a finite vector of real variables;
 %
 \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real matrices;
 %
 \item $b^i$, $c^i_l$, $c^i_u$ for $i \in \{ 0 \} \cup \Mcm$ and $f$ are, respectively, real $n$-vectors, real constants, real constants and real constant;
 %
 \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$;
 %
 \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$;
 %
 \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
\end{itemize}

@@

The QPLIB format is defined in Table~\ref{tab-qplib-format}.
The data is in free format (blanks separate values), but must occur in
the order given here. Any blank lines, or lines starting with any of the
characters !, \% or \# are ignored. Each term in the first column of
Table~\ref{tab-qplib-format} denotes a required value. Any strings beyond
those required on a given line will be regarded as comments and ignored.
Real values may either by in decimal or exponential formats; for the latter,
the exponent may be preceded by either the character \texttt{D} or \texttt{E},
e.g. \texttt{12.56D+2} or \texttt{12.56E+2}.

@@

$f$ & constant part of the objective function & \\ \hline

@@

$n^{b}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\

@@

$\infty$ & \lbox{value (real) for infinity for constraint or variable

@@

\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
would reduce flexibility. Always fix the default value for coefficients to 0
(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
default and then overwrite with the given entries, which may then be zero?)
and have no default bounds for variables. If you then also adapt the suggestion
to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
a left- and right-hand-side, then there is also no need to specify a value for infinity.
This would leave out ranged constraints - but are they really used? One could
do it like CPLEX and have an extra constraint sense for ranges and an extra
section to specify the range.}

@@

\begin{description}
\item [{[1]}]
\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}

The problem type is represented by a character string as one of the following:
\begin{description}
\item For continuous problems (i.e., all variables are continuous):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [LP]      a linear program,
\item [LPQC]    a linear program with quadratic constraints,
\item [BQP]     a bound-constrained quadratic program,
\item [QP]      a quadratic program, and
\item [QPQC]    a quadratic program with quadratic constraints.
\end{description}
\item For integer problems (i.e., all variables are integer valued):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [ILP]     an integer linear program,
\item [ILPQC]   an integer linear program with quadratic constraints,
\item [IBQP]    an integer bound-constrained quadratic program,
\item [IQP]     an integer quadratic program, and
\item [IQPQC]   an integer quadratic program with quadratic constraints.
\end{description}
\item For mixed-integer problems (i.e., there is a mix of continuous and
integer variables):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [MILP]    a mixed-integer linear program,
\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
\item [MIQP]    a mixed-integer quadratic program, and
\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
\end{description}
\todo[inline]{\small Stefan: I agree that leaving out information on
numerical properties of the matrices and vectors (convexity, etc.) is good here.
Such properties may also depend on tolerances, etc.
As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
\end{description}

@@

{\small
\begin{verbatim}
! ---------------
! example problem
! ---------------
MIPBAND   # problem name
MIQP      # problem is a mixed-integer quadratic program
Minimize  # minimize the objective function
3         # variables
2         # general linear constraints
5         # nonzeros in lower triangle of Q^0
1 1 2.0   5 lines of row & column index & value of nonzeros in lower triangle Q^0
2 1 -1.0  |
2 2 2.0   |
3 2 -1.0  |
3 3 2.0   |
-0.2      default value for entries in b_0
1         # non default entries in b_0
2 -0.4    1 lines of index & value of non-default values in b_0
0.0       value of f
4         # nonzeros in vectors b^i (i =0,...,m)
1 1 1.0   4 lines of row & column index & value of nonzeros in b^i (i =0,...,m)
1 2 1.0   |
2 1 1.0   |
2 3 1.0   |
1.0E+20   infinity
1.0       default value for entries in c_l
0         # non default entries in c_l
1.0E+20   default value for entries in c_u
0         # non default entries in c_u
0.0       default value for entries in l
0         # non default entries in l
1.0       default value for entries in u
1         # non default entries in u
2 2.0     1 line of non-default indices and values in u
0         default variable type is continuous
1         # non default variable types
3 1       variable 3 is integer
1.0       default value for initial values for x
0         # non default entries in x
0.0       default value for initial values for y
0         # non default entries in y
0.0       default value for initial values for z
0         # non default entries in z
0         # non default names for variables
0         # non default names for constraints
\end{verbatim}
}

>>>>

@@

The QPLIB format is defined in Table~\ref{tab-qplib-format}, with the notation of \S~\ref{sec:QPbasic}.

The data is in free format (blanks separate values), but must occur in
the order given here. Any blank lines, or lines starting with any of the
characters !, \% or \# are ignored. Each term in the first column of
Table~\ref{tab-qplib-format} denotes a required value. Any strings beyond
those required on a given line will be regarded as comments and ignored.
Real values may either by in decimal or exponential formats; for the latter,
the exponent may be preceded by either the character \texttt{D} or \texttt{E},
e.g. \texttt{12.56D+2} or \texttt{12.56E+2}.

@@

$q^0$ & constant part of the objective function & \\ \hline

@@

$n^{b^i}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\

@@

$c_{\infty}$ & \lbox{value (real) for infinity for constraint or variable

@@

%
%\todo[inline]{\small Stefan: I would suggest to make it simpler, even though that
%would reduce flexibility. Always fix the default value for coefficients to 0
%(otherwise, how you parse this? First initialize a n x n matrix with the non-zero
%default and then overwrite with the given entries, which may then be zero?)
%and have no default bounds for variables. If you then also adapt the suggestion
%to use a constraint sense ($\leq$,$\geq$,$=$) and a right-hand-side instead of
%a left- and right-hand-side, then there is also no need to specify a value for infinity.
%This would leave out ranged constraints - but are they really used? One could
%do it like CPLEX and have an extra constraint sense for ranges and an extra
%section to specify the range.}
%

@@

\begin{description}
\item [{[1]}]
%\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}

The problem type is represented by a character string as one of the following:
\begin{description}
\item For continuous problems (i.e., all variables are continuous):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [LP]      a linear program,
\item [LPQC]    a linear program with quadratic constraints,
\item [BQP]     a bound-constrained quadratic program,
\item [QP]      a quadratic program, and
\item [QPQC]    a quadratic program with quadratic constraints.
\end{description}
\item For integer problems (i.e., all variables are integer valued):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [ILP]     an integer linear program,
\item [ILPQC]   an integer linear program with quadratic constraints,
\item [IBQP]    an integer bound-constrained quadratic program,
\item [IQP]     an integer quadratic program, and
\item [IQPQC]   an integer quadratic program with quadratic constraints.
\end{description}
\item For mixed-integer problems (i.e., there is a mix of continuous and
integer variables):
\begin{description}[leftmargin=!,labelwidth=\widthof{\texttt{SMILQPC}}]
\item [MILP]    a mixed-integer linear program,
\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
\item [MIQP]    a mixed-integer quadratic program, and
\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
\end{description}
%\todo[inline]{\small Stefan: I agree that leaving out information on
%numerical properties of the matrices and vectors (convexity, etc.) is good here.
%Such properties may also depend on tolerances, etc.
%As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
%Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
\end{description}

@@

{\small
\begin{verbatim}
! ---------------
! example problem
! ---------------
MIPBAND   # problem name
MIQP      # problem is a mixed-integer quadratic program
Minimize  # minimize the objective function
3         # variables
2         # general linear constraints
5         # nonzeros in lower triangle of Q^0
1 1 2.0   5 lines of row & column index & value of nonzeros in lower triangle Q^0
2 1 -1.0  |
2 2 2.0   |
3 2 -1.0  |
3 3 2.0   |
-0.2      default value for entries in b_0
1         # non default entries in b_0
2 -0.4    1 lines of index & value of non-default values in b_0
0.0       value of q^0
4         # nonzeros in vectors b^i (i =0,...,m)
1 1 1.0   4 lines of row & column index & value of nonzeros in b^i (i =0,...,m)
1 2 1.0   |
2 1 1.0   |
2 3 1.0   |
1.0E+20   infinity
1.0       default value for entries in c_l
0         # non default entries in c_l
1.0E+20   default value for entries in c_u
0         # non default entries in c_u
0.0       default value for entries in l
0         # non default entries in l
1.0       default value for entries in u
1         # non default entries in u
2 2.0     1 line of non-default indices and values in u
0         default variable type is continuous
1         # non default variable types
3 1       variable 3 is integer
1.0       default value for initial values for x
0         # non default entries in x
0.0       default value for initial values for y
0         # non default entries in y
0.0       default value for initial values for z
0         # non default entries in z
0         # non default names for variables
0         # non default names for constraints
\end{verbatim}
}
